---
title: "Global_Gender_Inequality"
output: html_document
date: "2024-10-26"
---

```{r}
library(here)
library(readr)
library(sf)
library(dplyr)
library(countrycode)
```

#step 1.1: loading the inqeuality data

```{r}
global_gender_inequality <- read.csv(here("HDR23-24_Composite_indices_complete_time_series.csv")) %>%
  select(iso3, gii_2010, gii_2019)

```


#step 1.2: calculationg the difference between 2010 and 2019                                   
```{r}
gii20102019 <- global_gender_inequality %>%
  mutate(diff_gii = gii_2019 - gii_2010) %>%
  filter(!is.na(diff_gii))
```

#step 1.3: checking the data 
```{r}
head(gii20102019,10)

```

#step 2.1: loading the spatial data (GeoJSON)
```{r}
countries_GeoJSON <- st_read(here("World_Countries_(Generalized)_9029012925078512962.geojson"))
```

#step 2.2: adding iso3 column using the "countrycode" package
```{r}
countries_data <- countries_GeoJSON %>%
  mutate(iso3 = countrycode(ISO, origin = "iso2c", destination = "iso3c"))
```

#step 2.3: checking the data
```{r}
head(countries_data,10)

```

#step 2.4: too messy! let's clean up a bit

```{r}
countries_clean <- countries_data %>%
  select(COUNTRY, geometry, iso3)
```

#step 2.5: let's check the data again
```{r}
head(countries_clean, 10)
  #much better, nice and clean
```

#step 3.1: combining the inequality data to the spatial data using (left_join)
```{r}
merged <- countries_clean %>%
  left_join(gii20102019, by = "iso3")
```

#step 3.2: checking the combined data
```{r}
head(merged,10)
```

#step 4: creating a map for the combined data
```{r Map}
#Loading required packages
library(tmap)
library(tmaptools)
tmap_mode = "plot"

#Simplifying geometry for nicer borders (st_simplify wasn't working)
library(rmapshaper)
merged <- ms_simplify(merged, keep = 0.05)

#Map
tm_shape(merged)+
  tm_fill("diff_gii", 
              style="pretty",
              n=7,
              palette="-PuOr",
              midpoint=0,
              title="Change in GII", 
              alpha=0.9, 
              borders=FALSE, 
              textNA = "No Data",
              legend.hist = TRUE)+
    tm_borders(
        col="black",
        lty="solid",
        lwd=0.6)+
    tm_compass(position = c(0.005, 0.2), size=1.5)+
    tm_scale_bar(text.size=0.4,position=c(0.005, 0.1), width=0.2)+
    tm_layout(
        main.title.position = c("left", "top"),      
        main.title.size = 1.5,
        main.title="Change in UN Gender Inequality Index: 2010 to 2019",
        bg.color="skyblue",
        frame=TRUE,
        legend.outside=TRUE,
        outer.margins = c(0, 0, 0.1, 0))+
  tm_credits("Note: a decrease in GII constitutes an increase in gender equality.", size = 0.4, position=c(0.5, 0.02))


#Not sure if tmap is very difficult to use
#Or if it just doesn't work well on my laptop
#But leaving this here as I think I'll have a mental breakdown if I continue
```

